---
title: "simpleRegress"
author: "leia"
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


## make a simple regression model for age 
 
## Model LENGTH ~ AGE this is an example 

```{r}
AGEmodel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(AGEmodel)
summary.lm(AGEmodel)
```

## Model ~ LENGTH ~ HEADLEN this is an example 

```{r}
headlenght <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenght)
summary.lm(headlenght)
```
 
 
## Model ~ LENGTH ~ HEADWTH this is an example  

```{r}
headwidth <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwidth) 
summary.lm(headwidth)
```

## Model LENGTH ~ CHEST this an example 

```{r}
chestmodel <- lm(LENGTH~CHEST,data=bears) 
plot(LENGTH~CHEST,data=bears) 
abline(chestmodel) 
summary.lm(chestmodel)
```
 
 
## Model LENGTH ~ WEIGHT this is an example 
 
```{r}
weightmodel <- lm(LENGTH~WEIGHT,data=bears) 
plot(LENGTH~WEIGHT,data=bears) 
abline(weightmodel) 
summary.lm(weightmodel)
```


## conclusion

amange all the model with the largest rsquare was .8434 and the best one was head length. 